---
title: "NF Analysis"
author: Ghislain d'Entremont
output: html_document
---

# Data Processing

## Load Packages 

```{r message=FALSE, echo=FALSE}
library(signal)
library(tidyverse)
library(rstan)
library(bspec)
```

## Load and Tidy Data

```{r message=FALSE, echo=FALSE}
setwd("/Volumes/LaCie/Experiments/NF_data/Piloting")

raw = map_df(
  .x = list.files(
    pattern = "p99_raw"
  )
  , .f = function(file) {
    read_csv(
      file = file
      , col_names = FALSE
    )
  }
)

names(raw) = c('id', 'age', 'sex', 'hand', 'year', 'month', 'day', 'hour', 'minute', 'seconds', 'block' , 'block_cond', 'trial', 'cue_loc', 'iti', 'trial_stage', 'iteration', 'time', 'Ch1', 'Ch2', 'Ch3', 'Ch4', 'Ch5', 'Ch6', 'Ch7', 'Ch8', 'Ch9', 'Ch10', 'Ch11', 'Ch12', 'Ch13', 'Ch14')

raw$id = factor(raw$id)
raw$sex = factor(raw$sex, levels = c(1, 2), labels = c("male", "female"))
raw$hand = factor(raw$hand, levels = c(1, 2), labels = c("left", "right"))
raw$block = factor(raw$block, levels = c(1:5), labels = c("1", "2", "3", "4", "5"))
raw$block_cond = factor(raw$block_cond, levels = c(1, 2), labels = c("MI", "ME"))
raw$cue_loc = factor(raw$cue_loc, levels = c(1, 2), labels = c("left", "right"))
raw$trial_stage = factor(raw$trial_stage, levels = c(1:4), labels = c("iti", "fixation", "cue", "NF"))

raw_long = raw %>%
  gather(channel, signal, Ch1:Ch14, factor_key = TRUE)
```

##### Add adjusted time column that starts at zero for each trial for epochs later. We will be using estimated time based on the EEG system's sampling frequency (128 Hz). Therefore, we confirm that the estimated and actual timestamps don't differ substantially. Indeed, the timestamps are accurate to within 10 ms. 

```{r, echo=FALSE}
raw_long = raw_long %>%
  group_by(block_cond, block, trial, trial_stage, channel) %>%
  mutate(
    time_adj = time - min(time)
    , time_est = seq(0, (1/128)*(length(id)-1), by = 1/128)
    )

range(raw_long$time_est - raw_long$time_adj)
```

##### Acknowledge timing issue: there appears to be a time lag in the latter half of each block, specifically in the cue presentation trial stages. I'm unsure as to why right now. It is very consistent, so it may be a systematic error in the code. Specifically, the time seems to double...  

```{r, echo=FALSE}
raw_long %>%
  group_by(block_cond, block, trial, trial_stage) %>%
  summarize(max_time = max(time_est)) %>%
  filter(block != "1", trial_stage != "iti") %>%
  mutate(exp_time = (as.numeric(block) - 2) * 60 + as.numeric(trial)) %>%
  group_by(block_cond, exp_time, trial_stage) %>%
  ggplot()+
    geom_point(aes(x=exp_time, y=max_time))+
    facet_grid(block_cond~trial_stage)+
    ylab("Time of Trial Stage (s)")+
    xlab("Experiment Time (Trials)")
```

##### Filter data

```{r, echo=FALSE}
bu2 = butter(6, c(13,30)/(128/2), type = 'pass')

filtered = raw_long %>%
  group_by(block_cond, channel) %>%
  mutate(butter_filtered = signal::filter(bu2, signal)) %>%
  mutate(power = as.numeric(butter_filtered)^2)
```

##### Get rid of practice block

```{r, echo=FALSE}
filtered = filtered %>%
  dplyr::filter(block != "1")
```

##### Create laterality variable

```{r, echo=FALSE}
filtered = filtered %>%
  dplyr::filter(channel == "Ch8" | channel == "Ch12") %>%
  mutate(laterality = ifelse(
    (cue_loc == "right" & channel == "Ch8") | (cue_loc == "left" & channel == "Ch12")
      , "contra"
      , "ipsi"
    )
  ) 
```


# ERD Timecourses

## (Log of Average/Average of log of Power) over Moving Average Baseline

##### Get moving average baseline. For now, I use make a window of size 1, and therefore get a different baseline for each trial. This actually ensures that the ERDs are relative to zero. 

```{r, echo=FALSE}
baselines = filtered %>%
  dplyr::filter(trial_stage == "fixation") %>%
  group_by(block_cond, block, trial) %>%
  summarize(baseline = mean(power))

# baselines = baselines %>%
#   filter(!is.na(baseline))

WINDOW_SIZE = 1

mavgs = NULL
for (i in 1:nrow(baselines)) {

  if (baselines$block[i] == 2 & baselines$trial[i] == 1) {
    window = NULL
  }

  window = c(baselines$baseline[i], window)

  if (length(window) > WINDOW_SIZE) {
      window_40 = window[1:WINDOW_SIZE]
  } else {
    window_40 = window
  }

  mavg = mean(window_40)
  mavgs = c(mavgs, mavg)
}

mavgs_tibb = tibble(block_cond = baselines$block_cond,  block = baselines$block, trial = baselines$trial, mavg = mavgs)
```

##### Get relative power

```{r, echo=FALSE}
for_ERD = filtered %>%
  dplyr::filter(trial_stage != "iti") %>%
  group_by(block_cond, block, trial, laterality) %>%
  mutate(
  time_est_ERD = seq(0, (1/128)*(length(id)-1), by = 1/128)
  ) %>%
  left_join(mavgs_tibb) %>%
  mutate(
    relative_power = (power/mavg)
    , log_relative_power = log(relative_power)
  )
```

### Log of Average over Moving Average Baseline

```{r, echo=FALSE}
for_ERD %>%
  group_by(block_cond, time_est_ERD, laterality) %>%
  summarize(avg_relative_power = mean(relative_power, na.rm = T)) %>%
  group_by(block_cond, laterality) %>%
  mutate(sma_relative_power = stats::filter(avg_relative_power, rep(1/128, 128), sides=2)) %>%
  mutate(log_sma_relative_power = log(sma_relative_power)) %>%
  mutate(log_avg_relative_power = log(avg_relative_power)) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=log_avg_relative_power, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=log_sma_relative_power, color = laterality))+
    facet_grid(.~block_cond)+
    xlab("Trial Time (s)")+
    ylab("Log of Average Relative Power")+
    ggtitle("All Trials") +
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-1, 1))  # no avoid outlying noise at the end

for_ERD %>%
  dplyr::filter(trial <= 30) %>%
  group_by(block_cond, time_est_ERD, laterality) %>%
  summarize(avg_relative_power = mean(relative_power, na.rm = T)) %>%
  group_by(block_cond, laterality) %>%
  mutate(sma_relative_power = stats::filter(avg_relative_power, rep(1/128, 128), sides=2)) %>%
  mutate(log_sma_relative_power = log(sma_relative_power)) %>%
  mutate(log_avg_relative_power = log(avg_relative_power)) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=log_avg_relative_power, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=log_sma_relative_power, color = laterality))+
    facet_grid(.~block_cond)+
    xlab("Trial Time (s)")+
    ylab("Log of Average Relative Power")+
    ggtitle("First 30 Trials") +
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-1, 1))  # no avoid outlying noise at the end
```

### Average of Log over Moving Average Baseline

##### Taking the average of the logs spits out values centered around -2 on the y-axis. This is not ideal. I run a miniature simulation below to see how this shift may work itself out mathematically. Even if relative power over the course of the baseline period is on average 1 (as it should be because that's how baseline was defined for each trial), you can still get negative values if you take the log of each data point. 

```{r, echo=FALSE}
for_ERD %>%
  group_by(block_cond, time_est_ERD, laterality) %>%
  summarize(avg_log_relative_power = mean(log_relative_power, na.rm = T)) %>%
  group_by(block_cond, laterality) %>%
  mutate(sma_log_relative_power = stats::filter(avg_log_relative_power, rep(1/128, 128), sides=2)) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=avg_log_relative_power, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=sma_log_relative_power, color = laterality))+
    facet_grid(.~block_cond)+
    xlab("Trial Time (s)")+
    ylab("Average of Log of Relative Power")+
    ggtitle("All Trials") +
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-4, 0))

for_ERD %>%
  filter(trial <= 30) %>%
  group_by(block_cond, time_est_ERD, laterality) %>%
  summarize(avg_log_relative_power = mean(log_relative_power, na.rm = T)) %>%
  group_by(block_cond, laterality) %>%
  mutate(sma_log_relative_power = stats::filter(avg_log_relative_power, rep(1/128, 128), sides=2)) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=avg_log_relative_power, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=sma_log_relative_power, color = laterality))+
    facet_grid(.~block_cond)+
    xlab("Trial Time (s)")+
    ylab("Average of Log of Relative Power")+
    ggtitle("First 30 Trials") +
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-4, 0))

mean(log(runif(1000000, 1-1, 1+1)/1))

mean(log(rlnorm(1000000, 0 - 1/2)/1))
```

## Percent ERD after Averaging across Trials (Zich & Pfurtscheller)

##### Average power samples across trial, get SMA and baseline, and plot % ERD

```{r, echo=FALSE}
SMA = filtered %>%
  dplyr::filter(trial_stage != "iti") %>%
  group_by(block_cond, block, trial, laterality) %>%
  mutate(
    time_est_ERD = seq(0, (1/128)*(length(id)-1), by = 1/128)
    ) %>%
  group_by(block_cond, time_est_ERD, laterality) %>% 
  summarize(avg_power = mean(power)) %>%
  # left_join(avg_baselines) %>%
  group_by(block_cond, laterality) %>%
  mutate(sma_power = stats::filter(avg_power, rep(1/128, 128), sides=2))

SMA_baseline = SMA %>%
  dplyr::filter(0 <= time_est_ERD & time_est_ERD < 2.00) %>%
  group_by(block_cond) %>%
  summarize(avg_baseline = mean(as.numeric(sma_power), na.rm = T))

SMA %>%
  left_join(SMA_baseline) %>%
  mutate(sma_percent_ERD = ((as.numeric(sma_power) - avg_baseline)/avg_baseline) * 100)%>%
  mutate(percent_ERD = ((avg_power - avg_baseline)/avg_baseline) * 100) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=percent_ERD, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=sma_percent_ERD, color = laterality))+
    facet_grid(.~block_cond)+
    xlab("Trial Time (s)")+
    ylab("Percent ERD")+
    ggtitle("All Trials")+
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-50,50))


SMA_30 = filtered %>%
  dplyr::filter(trial_stage != "iti", trial <= 30) %>%
  group_by(block_cond, block, trial, laterality) %>%
  mutate(
    time_est_ERD = seq(0, (1/128)*(length(id)-1), by = 1/128)
    ) %>%
  group_by(block_cond, time_est_ERD, laterality) %>% 
  summarize(avg_power = mean(power)) %>%
  # left_join(avg_baselines) %>%
  group_by(block_cond, laterality) %>%
  mutate(sma_power = stats::filter(avg_power, rep(1/128, 128), sides=2))

SMA_baseline_30 = SMA_30 %>%
  dplyr::filter(0 <= time_est_ERD & time_est_ERD < 2.00) %>%
  group_by(block_cond) %>%
  summarize(avg_baseline = mean(as.numeric(sma_power), na.rm = T))

SMA_30 %>%
  left_join(SMA_baseline_30) %>%
  mutate(sma_percent_ERD = ((as.numeric(sma_power) - avg_baseline)/avg_baseline) * 100)%>%
  mutate(percent_ERD = ((avg_power - avg_baseline)/avg_baseline) * 100) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=percent_ERD, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=sma_percent_ERD, color = laterality))+
    facet_grid(.~block_cond)+
    xlab("Trial Time (s)")+
    ylab("Percent ERD")+
    ggtitle("First 30 Trials")+
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-50,50))
```

# Laterality Index

##### Get mean LI per trial. NOTE: the average baselines used herein are still based on single trials (i.e. 'moving average' of window of a single trial).

```{r, echo=FALSE}
LIs = filtered %>%
  dplyr::filter(trial_stage == "NF") %>%
  group_by(block_cond, block, trial, laterality) %>%
  summarize(power = mean(power)) %>%
  left_join(mavgs_tibb) %>%
  spread(key = laterality, power) %>%
  mutate(LI = log(ipsi/mavg) - log(contra/mavg))
```

##### Plot LI as a function of experiment time

```{r, echo=FALSE}
LIs %>%
  mutate(exp_time = (as.numeric(block) - 2) * 60 + as.numeric(trial)) %>%
  ggplot()+
    geom_line(aes(x=exp_time, y=LI))+
    facet_grid(block_cond ~ .)+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylab("Laterality Index\n(Difference in Log of Power Ratio)")+
    xlab("Experiment Time (Trials)")
```



