---
title: "NF Analysis"
author: Ghislain d'Entremont
output: html_document
---

# Data Processing

### Load Packages 

```{r message=FALSE, echo=FALSE}
library(signal)
library(tidyverse)
library(rstan)
library(bspec)
```

### Load and Tidy Data

```{r message=FALSE}
setwd("/Volumes/LaCie/Experiments/NF_data/Piloting/chelsey_pilot")

raw = map_df(
  .x = list.files(
    pattern = "raw"
  )
  , .f = function(file) {
    read_csv(
      file = file
      , col_names = FALSE
    )
  }
)

# matlab mappings:
# [id age sex hand year month day hour minute seconds feedback task block trial cue_loc_idx iti trial_stage iteration]

names(raw) = c('id', 'age', 'sex', 'hand', 'year', 'month', 'day', 'hour', 'minute', 'seconds', 'feedback', 'task', 'block', 'trial', 'cue_loc', 'iti', 'trial_stage', 'iteration', 'time', 'Ch1', 'Ch2', 'Ch3', 'Ch4', 'Ch5', 'Ch6', 'Ch7', 'Ch8', 'Ch9', 'Ch10', 'Ch11', 'Ch12', 'Ch13', 'Ch14')

raw$id = factor(raw$id)
raw$sex = factor(raw$sex, levels = c(1, 2), labels = c("male", "female"))
raw$hand = factor(raw$hand, levels = c(1, 2), labels = c("left", "right"))
raw$block = factor(raw$block, levels = c(1:5), labels = c("1", "2", "3", "4", "5"))
raw$task = factor(raw$task, levels = c(1, 2), labels = c("MI", "ME"))
raw$feedback = factor(raw$feedback, levels = c(1, 2), labels = c("C", "I"))
raw$cue_loc = factor(raw$cue_loc, levels = c(1, 2), labels = c("left", "right"))
raw$trial_stage = factor(raw$trial_stage, levels = c(1:5), labels = c("iti", "fixation", "cue", "NF", "inter"))

# summary(raw)

raw_long = raw %>%
  gather(channel, signal, Ch1:Ch14, factor_key = TRUE)

summary(raw_long)

# delete original df
rm(raw)
```

##### Time Stamps

Add adjusted time column that starts at zero for each trial for epochs later. We will be using estimated time based on the EEG system's sampling frequency (128 Hz). Therefore, we confirm that the estimated and actual timestamps don't differ substantially. Indeed, the timestamps are accurate to within 10 ms. 

```{r, echo=FALSE}
raw_long = raw_long %>%
  group_by(task, block, trial, trial_stage, channel) %>%
  mutate(
    time_adj = time - min(time)
    , time_est = seq(0, (1/128)*(length(id)-1), by = 1/128)
    )

range(raw_long$time_est - raw_long$time_adj)
```

##### Trial Stage Timing

```{r, echo=FALSE}
raw_long %>%
  group_by(task, block, trial, trial_stage) %>%
  summarize(max_time = max(time_est)) %>%
  filter(block != "1", trial_stage != "iti") %>%
  mutate(exp_time = (as.numeric(block) - 2) * 60 + as.numeric(trial)) %>%
  group_by(task, exp_time, trial_stage) %>%
  ggplot()+
    geom_point(aes(x=exp_time, y=max_time))+
    facet_grid(task~trial_stage)+
    ylab("Time of Trial Stage (s)")+
    xlab("Experiment Time (Trials)")
```

##### Filter data

```{r, echo=FALSE}
bu = butter(6, c(13,30)/(128/2), type = 'pass')

filtered = raw_long %>%
  group_by(task, channel) %>%
  mutate(butter_filtered = signal::filtfilt(bu, signal)) %>%
  mutate(signal_squared = butter_filtered^2)

filtered %>%
  dplyr::filter(block == 2, trial <= 5) %>%
  ggplot(aes(y = butter_filtered, x = time))+
    geom_line()+
    facet_grid(channel ~task, scales = "free")+
    ggtitle("Partial Filtered Experiment Data")+
    xlab("Time (s)")+
    ylab("Signal (uV)")

# remove long format df
rm(raw_long)
```

##### Get rid of practice block

```{r, echo=FALSE}
filtered = filtered %>%
  dplyr::filter(block != "1")
```

##### Create laterality variable

```{r, echo=FALSE}
filtered = filtered %>%
  dplyr::filter(channel == "Ch8" | channel == "Ch12") %>%
  mutate(laterality = ifelse(
    (cue_loc == "right" & channel == "Ch8") | (cue_loc == "left" & channel == "Ch12")
      , "contra"
      , "ipsi"
    )
  ) 
```

# ERD Timecourses

### Moving Average Baseline

I use make a window of size 1, and therefore get a different baseline for each trial. This actually ensures that the ERDs are relative to zero. 

```{r, echo=FALSE}
baselines = filtered %>%
  dplyr::filter(trial_stage == "fixation") %>%
  group_by(task, block, trial) %>%
  summarize(baseline = mean(signal_squared))

# baselines = baselines %>%
#   filter(!is.na(baseline))

WINDOW_SIZE = 1

mavgs = NULL
for (i in 1:nrow(baselines)) {

  if (baselines$block[i] == 2 & baselines$trial[i] == 1) {
    window = NULL
  }

  window = c(baselines$baseline[i], window)

  if (length(window) > WINDOW_SIZE) {
      window_40 = window[1:WINDOW_SIZE]
  } else {
    window_40 = window
  }

  mavg = mean(window_40)
  mavgs = c(mavgs, mavg)
}

mavgs_tibb = tibble(task = baselines$task,  block = baselines$block, trial = baselines$trial, mavg = mavgs)

# look for outliying baseline trials
cutoff = 50
mavgs_tibb %>%
  ggplot()+
    geom_histogram(aes(x=mavgs), bins = 1000)+
    geom_vline(xintercept = cutoff, linetype = "dotted", color = "red")

# identify them 
mavgs_tibb %>%
  mutate(
    exclude = ifelse(mavg >= cutoff, TRUE, FALSE)
    ) -> mavgs_tibb

# examine them
mavgs_tibb %>%
  filter(exclude)
```

### Get relative power

We exclude trials for which the baselines were bad. We then examine the distribution of relative power and log relative power. First, it seems as though the continuous values of log relative power are not normally distributed. They are skewed to the left. Second, the distibution of relative power is skewed to the right. There is quite an extrordinary range of relative power values. From virtually zero to 4000.

Currently, we make no exclusions based on relative power or log relative power.

```{r, echo=FALSE}
for_ERD = filtered %>%
  dplyr::filter(trial_stage != "iti") %>%
  group_by(task, block, trial, laterality) %>%
  mutate(
  time_est_ERD = seq(0, (1/128)*(length(id)-1), by = 1/128)
  ) %>%
  left_join(mavgs_tibb) %>%
  filter(exclude == FALSE) %>%
  mutate(
    relative_power = (signal_squared/mavg)
    , log_relative_power = log(relative_power)
  )

# look for outlying log relative power values
for_ERD %>%
  ggplot()+
    geom_histogram(aes(x=log_relative_power), bins = 1000)

sort(for_ERD$log_relative_power)[c(1:10, (nrow(for_ERD)-10):nrow(for_ERD))]

# look for outlying relative power values
for_ERD %>%
  ggplot()+
    geom_histogram(aes(x=relative_power), bins = 1000)+
    xlim(c(0, 20))

sort(for_ERD$relative_power)[c(1:10, (nrow(for_ERD)-10):nrow(for_ERD))]
```

### Log of Average over Moving Average Baseline

It is difficult to dicern any clear pattern from the data.

```{r, echo=FALSE}
for_ERD %>%
  group_by(task, time_est_ERD, laterality) %>%
  summarize(avg_relative_power = mean(relative_power, na.rm = T)) %>%
  group_by(task, laterality) %>%
  mutate(sma_relative_power = stats::filter(avg_relative_power, rep(1/128, 128), sides=2)) %>%
  mutate(log_sma_relative_power = log(sma_relative_power)) %>%
  mutate(log_avg_relative_power = log(avg_relative_power)) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=log_avg_relative_power, color = laterality), alpha = 0.3)+
    geom_line(aes(x=time_est_ERD, y=log_sma_relative_power, color = laterality))+
    facet_grid(.~task)+
    xlab("Trial Time (s)")+
    ylab("Log of Average Relative Power")+
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    # geom_vline(xintercept = 2 + 1.25 + 5, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+ 
    xlim(c(-0.25, 8.5))+
    ylim(c(-2.5, 5))
```

### Percent ERD after Averaging across Trials (Zich & Pfurtscheller)

Average power samples across trial, get SMA and baseline, and plot % ERD... 

The data are even less clear here.

```{r, echo=FALSE}
SMA = filtered %>%
  dplyr::filter(trial_stage != "iti") %>%
  group_by(task, block, trial, laterality) %>%
  mutate(
    time_est_ERD = seq(0, (1/128)*(length(id)-1), by = 1/128)
    ) %>%
  group_by(task, time_est_ERD, laterality) %>% 
  summarize(avg_power = mean(signal_squared)) %>%
  # left_join(avg_baselines) %>%
  group_by(task, laterality) %>%
  mutate(sma_power = stats::filter(avg_power, rep(1/128, 128), sides=2))

SMA_baseline = SMA %>%
  dplyr::filter(0 <= time_est_ERD & time_est_ERD < 2.00) %>%
  group_by(task) %>%
  summarize(avg_baseline = mean(as.numeric(sma_power), na.rm = T))

SMA %>%
  left_join(SMA_baseline) %>%
  mutate(sma_percent_ERD = ((as.numeric(sma_power) - avg_baseline)/avg_baseline) * 100)%>%
  mutate(percent_ERD = ((avg_power - avg_baseline)/avg_baseline) * 100) %>%
  ggplot()+
    geom_line(aes(x=time_est_ERD, y=percent_ERD, color = laterality), alpha = 0.2)+
    geom_line(aes(x=time_est_ERD, y=sma_percent_ERD, color = laterality))+
    facet_grid(.~task)+
    xlab("Trial Time (s)")+
    ylab("Percent ERD")+
    ggtitle("All Trials")+
    geom_vline(xintercept = 2, linetype = "dashed")+
    geom_vline(xintercept = 2 + 1.25, linetype = "dashed")+
    geom_hline(yintercept=0, linetype = "dashed")+
    ylim(c(-50,50))
```


