---
title: "Checking MATLAB Output"
author: Ghislain d'Entremont
output: html_document
---

# Confirm Signal Quality and Validity 

### Load Packages 

```{r message=FALSE}
library(signal)
library(tidyverse)
library(rstan)
library(bspec)
```

### Load and Tidy Data

##### Raw Data

```{r message=FALSE}
setwd("/Volumes/LaCie/Experiments/NF_data/Piloting/chelsey_pilot")

raw = map_df(
  .x = list.files(
    pattern = "raw"
  )
  , .f = function(file) {
    read_csv(
      file = file
      , col_names = FALSE
    )
  }
)

# matlab mappings:
# [id age sex hand year month day hour minute seconds feedback task block trial cue_loc_idx iti trial_stage iteration]

names(raw) = c('id', 'age', 'sex', 'hand', 'year', 'month', 'day', 'hour', 'minute', 'seconds', 'feedback', 'task', 'block', 'trial', 'cue_loc', 'iti', 'trial_stage', 'iteration', 'time', 'Ch1', 'Ch2', 'Ch3', 'Ch4', 'Ch5', 'Ch6', 'Ch7', 'Ch8', 'Ch9', 'Ch10', 'Ch11', 'Ch12', 'Ch13', 'Ch14')

raw$id = factor(raw$id)
raw$sex = factor(raw$sex, levels = c(1, 2), labels = c("male", "female"))
raw$hand = factor(raw$hand, levels = c(1, 2), labels = c("left", "right"))
raw$block = factor(raw$block, levels = c(1:5), labels = c("1", "2", "3", "4", "5"))
raw$task = factor(raw$task, levels = c(1, 2), labels = c("MI", "ME"))
raw$feedback = factor(raw$feedback, levels = c(1, 2), labels = c("C", "I"))
raw$cue_loc = factor(raw$cue_loc, levels = c(1, 2), labels = c("left", "right"))
raw$trial_stage = factor(raw$trial_stage, levels = c(1:5), labels = c("iti", "fixation", "cue", "NF", "inter"))

# summary(raw)

raw_long = raw %>%
  gather(channel, signal, Ch1:Ch14, factor_key = TRUE)

summary(raw_long)

# delete original df
rm(raw)
```



##### Presentation Data

```{r message=FALSE}
setwd("/Volumes/LaCie/Experiments/NF_data/Piloting/chelsey_pilot")

pres = map_df(
  .x = list.files(
    pattern = "pres"
  )
  , .f = function(file) {
    read_csv(
      file = file
      , col_names = FALSE
    )
  }
)

# [id age sex hand year month day hour minute seconds feedback task block trial cue_loc_idx iti trial_stage iteration power_rest_mavg mean(Pxx(:,1)) mean(Pxx(:,2)) LI log_ERS_ipsi log_ERS_contra LOG_ERS_SCALE]
names(pres) = c('id', 'age', 'sex', 'hand', 'year', 'month', 'day', 'hour', 'minute', 'seconds', 'feedback', 'task', 'block', 'trial', 'cue_loc', 'iti', 'trial_stage', 'iteration', 'baseline_power_mavg', 'C3', 'C4', 'LI', 'log_ERS_ipsi', 'log_ERS_contra', 'disp_multi_factor') 

pres$id = factor(pres$id)
pres$sex = factor(pres$sex, levels = c(1, 2), labels = c("male", "female"))
pres$hand = factor(pres$hand, levels = c(1, 2), labels = c("left", "right"))
pres$block = factor(pres$block, levels = c(1:5), labels = c("1", "2", "3", "4", "5"))
pres$feedback = factor(pres$feedback, levels = c(1, 2), labels = c("C", "I")) 
pres$task = factor(pres$task, levels = c(1, 2), labels = c("MI", "ME"))
pres$cue_loc = factor(pres$cue_loc, levels = c(1, 2), labels = c("left", "right"))
pres$trial_stage = factor(pres$trial_stage, levels = c(1:5), labels = c("iti", "fixation", "cue", "NF", "inter"))

# summary(pres)

pres_long = pres %>%
  gather(channel, signal, c(C3, C4), factor_key = TRUE)

summary(pres_long)

# delete original df
rm(pres)
```

### Look at Raw Data

##### Check Time Stamps

We verify that time stamps are accurate. Indeed, the difference between the time stamps and the expected time points (based on sampling frequency) vary closely around 0. We confirm that there are no outliying discrepencies among the time stamps by examining the range, which is within 10 ms in absolute value for both block conditions.

```{r}
raw_long = raw_long %>%
  group_by(task, block, trial, trial_stage, channel) %>%
  mutate(
    time_adj = time - min(time)
    , time_est = seq(0, (1/128)*(length(id)-1), by = 1/128)
    )

raw_long %>% 
  dplyr::filter(block == 2, trial <= 5) %>%
  ggplot(aes(x = time, y = time_est - time_adj))+
    geom_point(size = 0.1)+
    facet_grid(.~task, scales = "free")+
    ggtitle("Timing Error through Time (first five trials)")+
    xlab("Time (s)")+
    ylab("Estimated - Recorded Times (s)")

range(raw_long$time_est - raw_long$time_adj)
```

##### Check for Timing Issues

We check for any time lags or other timing issues.

```{r}
raw_long %>%
  group_by(task, block, trial, trial_stage) %>%
  summarize(max_time = max(time_est)) %>%
  dplyr::filter(block != "1", trial_stage != "iti") %>%
  mutate(time_stage = as.numeric(block) * 60 + as.numeric(trial)) %>%
  group_by(task, time_stage, trial_stage) %>%
  ggplot()+
    geom_point(aes(x=time_stage, y=max_time))+
    facet_grid(task~trial_stage)
```


##### Raw waveforms and Epochs.

```{r}
raw_long %>%
  dplyr::filter(block == 2, trial <= 5) %>%
  ggplot(aes(y = signal, x = time))+
    geom_line()+
    facet_grid(channel ~ task, scales = "free")+
    ggtitle("Partial Raw Experiment Data")+
    xlab("Time (s)")+
    ylab("Signal (uV)")

raw_long %>%
  dplyr::filter(trial_stage == "NF", block == 2, trial <= 5) %>%
  group_by(task, channel, time_adj) %>%
  summarize(avg_signal = mean(signal)) %>%
  ggplot(aes(y = avg_signal, x = time_adj))+
    geom_line()+
    facet_grid(channel ~ task)+
    ggtitle("Partial NF Trial Grand Average")+
    xlab("Time (s)")+
    ylab("Signal (uV)")
```

### Process data 

##### Filter & Square

We look to create power values from raw data. For each task and each channel we will (1) apply a steep forward-backward butterworth filter and (2) square the signal to get continous power values.

```{r}
bu = butter(6, c(13,30)/(128/2), type = 'pass')
# no = butter(2, c(59,61)/(128/2), type = 'stop')

raw_long_fil = raw_long %>%
  group_by(task, channel) %>%
  mutate(butter_filtered = signal::filtfilt(bu, signal)) %>%
  # mutate(and_notch_filtered = signal::filtfilt(no, butter_filtered)) %>%
  mutate(signal_squared = butter_filtered^2)

raw_long_fil %>%
  dplyr::filter(block == 2, trial <= 5) %>%
  ggplot(aes(y = butter_filtered, x = time))+
  # ggplot(aes(y = and_notch_filtered, x = time))+
    geom_line()+
    facet_grid(channel ~task, scales = "free")+
    ggtitle("Partial Filtered Experiment Data")+
    xlab("Time (s)")+
    ylab("Signal (uV)")

raw_long_pwr = raw_long_fil %>%
  dplyr::filter(channel == "Ch8" || channel == "Ch12") %>%
  group_by(task, block, trial, trial_stage, cue_loc, channel) %>%
  summarize(power = mean(signal_squared))

# remove long format df
rm(raw_long)

# raw_long_pwr = raw_long_fil %>%
#   dplyr::filter(channel == "Ch8" || channel == "Ch12") %>%
#   group_by(task, block, trial, trial_stage, cue_loc, channel) %>%
#   summarize(power = mean(welchPSD(ts(and_notch_filtered), seglength = 128)[["power"]][14:31]))
```

<!-- ##### We confirm that welchPSD is doing what we want it to by visualizing the PSD of some decently sized middle chunk of the data for each channel of interest, using empirical power and welch's power. -->

<!-- # ```{r} -->
<!-- # raw_long_fil_Ch8 = raw_long_fil %>% -->
<!-- #   dplyr::filter(channel == "Ch8", task == "MI") -->
<!-- # -->
<!-- # len = nrow(raw_long_fil_Ch8) -->
<!-- # for_pwr = raw_long_fil_Ch8$and_notch_filtered[(len/2):(len/2+len/100)] -->
<!-- # -->
<!-- # spec_me = empiricalSpectrum(ts(for_pwr)) -->
<!-- # -->
<!-- # spec_me2 = welchPSD(ts(for_pwr), seglength = 128) -->
<!-- # -->
<!-- # plot(spec_me$freq*128, spec_me$power,  type="l", xlab="Frequency (Hz)", ylab="Power", main = "Channel 8") -->
<!-- # lines(spec_me2$frequency*128, spec_me2$power, col = "red") -->
<!-- # -->
<!-- # plot(spec_me$freq*128, spec_me$power, log = "y",  type="l", xlab="Frequency (Hz)", ylab="Power", main = "Channel 8") -->
<!-- # lines(spec_me2$frequency*128, spec_me2$power, col = "red") -->
<!-- # -->
<!-- # -->
<!-- # raw_long_fil_Ch12 = raw_long_fil %>% -->
<!-- #   dplyr::filter(channel == "Ch12", task == "MI") -->
<!-- # -->
<!-- # len = nrow(raw_long_fil_Ch12) -->
<!-- # for_pwr = raw_long_fil_Ch12$and_notch_filtered[(len/2):(len/2+len/100)] -->
<!-- # -->
<!-- # spec_me = empiricalSpectrum(ts(for_pwr)) -->
<!-- # -->
<!-- # spec_me2 = welchPSD(ts(for_pwr), seglength = 128) -->
<!-- # -->
<!-- # plot(spec_me$freq*128, spec_me$power,  type="l", xlab="Frequency (Hz)", ylab="Power", main = "Channel 12") -->
<!-- # lines(spec_me2$frequency*128, spec_me2$power, col = "red") -->
<!-- # -->
<!-- # plot(spec_me$freq*128, spec_me$power, log = "y",  type="l", xlab="Frequency (Hz)", ylab="Power", main = "Channel 12") -->
<!-- # lines(spec_me2$frequency*128, spec_me2$power, col = "red") -->
<!-- ``` -->

### Compare Presentation Values with Offline Calculations

<!-- ##### Let's compare the calculated power values to those that were computed in realtime in MATLAB. We see a very strong positive relationship between these power values for each channel. The discrency appears to lie in the magnitudes of the values, with the MATLAB computed power values being quite a bit smaller. The power values computed in real time were taken from smaller time windows and averaged (moving averages were themselves averaged), whereas those computed above were done for each trial stage. Both types of power estimates, however, were computed using welch's PSD. -->

##### Comparing Mean Trial Power

Power values appear reasonably, but not totally, related. The discrepency would come from the fact that realtime power values are calculated based on a data buffer - at each iteration, the data from the buffer are filtered and squared. The offline power values are calculated by first filtering the entire continuous signal, and then taking the average for a given 'trial stage' of the squared signal. 

```{r}
pwr_cal = raw_long_pwr %>%
  dplyr::filter(block != "1", trial_stage != "inter") %>%
  group_by(task, block, trial, trial_stage, channel) %>%
  summarize(avg_cal = mean(power)) 

pwr_pres = pres_long %>%
  dplyr::filter(block != "1", trial_stage != "inter") %>%
  group_by(task, block, trial, trial_stage, channel) %>%
  summarize(avg_real = mean(signal)) 

pwr_compare = bind_cols(pwr_cal, pwr_pres[,6])

# iti
pwr_compare %>% 
  dplyr::filter(trial_stage == "iti") %>%
  ggplot(aes(x = avg_cal, y = avg_real))+
    geom_point()+
    geom_smooth()+ 
    facet_grid(channel~task)+
    ggtitle("iti")+
    xlab("Calculated Power")+
    ylab("Realtime Power")+
    xlim(c(0, 500))+
    ylim(c(0, 500))+
    coord_fixed()

# fixation
pwr_compare %>% 
  dplyr::filter(trial_stage == "fixation") %>%
  ggplot(aes(x = avg_cal, y = avg_real))+
    geom_point()+
    geom_smooth()+ 
    facet_grid(channel~task)+
    ggtitle("fixation")+
    xlab("Calculated Power")+
    ylab("Realtime Power")+
    xlim(c(0, 500))+
    ylim(c(0, 500))+
    coord_fixed()

# cue
pwr_compare %>% 
  dplyr::filter(trial_stage == "cue") %>%
  ggplot(aes(x = avg_cal, y = avg_real))+
    geom_point()+
    geom_smooth()+ 
    facet_grid(channel~task)+
    ggtitle("cue")+
    xlab("Calculated Power")+
    ylab("Realtime Power")+
    xlim(c(0, 500))+
    ylim(c(0, 500))+
    coord_fixed()

# NF
pwr_compare %>% 
  dplyr::filter(trial_stage == "NF") %>%
  ggplot(aes(x = avg_cal, y = avg_real))+
    geom_point()+
    geom_smooth()+ 
    facet_grid(channel~task)+
    ggtitle("NF")+
    xlab("Calculated Power")+
    ylab("Realtime Power")+
    xlim(c(0, 500))+
    ylim(c(0, 500))+
    coord_fixed()

pwr_compare[,-c(2,3,4)] %>%
  group_by(task, channel) %>%
  summarize(corr = cor(avg_cal, avg_real, use = "complete"))
```

<!-- ##### Now, we can compute baseline power values from our calculated power values and compare those with those of the realtime MATAB output. The hardly seems to be a relationship in the ME condition. I'm unsure as to why right now... We've already shown that the power values correlate with one another. The moving baseline, if computed properly, should related in some way. Interestingly, the moving average baselines calculated offline and in real time match up well in the MI condition.  -->

##### Compare Trial-wise Baseline Power

The baseline power calculations are virtually identical. This is expected as neither the online nor the offline baseline calculations involved the use of a data buffer. 

```{r}
baseline_pwr_cal = pwr_cal %>%
  dplyr::filter(trial_stage == "fixation") %>%
  group_by(task, block, trial) %>% 
  summarize(baseline = mean(avg_cal))

mavgs = NULL
for (i in 1:nrow(baseline_pwr_cal)) {
  
  if (baseline_pwr_cal$block[i] == 2 & baseline_pwr_cal$trial[i] == 1) {
    window = NULL
  }
  
  window = c(baseline_pwr_cal$baseline[i], window)
  
  if (length(window) > 1) {
      window_40 = window[1:1]
  } else {
    window_40 = window
  }
  
  mavg = mean(window_40)
  mavgs = c(mavgs, mavg)
}

mavg_pwr_cal = tibble(task = baseline_pwr_cal$task,  block = baseline_pwr_cal$block, trial = baseline_pwr_cal$trial, cal_mavg = mavgs) 
  
mavg_pwr_pres = pres_long %>% 
  dplyr::filter(trial_stage == "NF", block != "1") %>%
  group_by(task, block, trial) %>%
  summarize(real_mavg = mean(baseline_power_mavg)) 

mavg_compare = left_join(mavg_pwr_cal, mavg_pwr_pres)  # this gets rid of NA

mavg_compare %>%
  ggplot(aes(x = cal_mavg, y = real_mavg))+
    geom_point()+
    geom_smooth()+ 
    facet_grid(.~task)+
    ggtitle("Moving Average Baseline Power")+
    xlab("Calculated Baseline")+
    ylab("Realtime Basline")+
    coord_fixed()

mavg_compare %>%
  group_by(task) %>%
  summarize(corr = cor(cal_mavg, real_mavg, use = "complete"))
```

##### Compare Log Relative Power

The log are almost perfectly related. The correlation comes out to be very large. This is true for both tasks. 

```{r}
rel_pwr_cal = raw_long_pwr %>%
  dplyr::filter(trial_stage == "NF", block != "1") %>%
  left_join(mavg_pwr_cal) %>%
  spread(key = channel, power) %>%
  mutate(
    log_ERS_ipsi_cal = ifelse(
      cue_loc == "right"
      , log(Ch12/cal_mavg)
      , log(Ch8/cal_mavg)
    )
    , log_ERS_contra_cal = ifelse(
      cue_loc == "right"
      , log(Ch8/cal_mavg) 
      , log(Ch12/cal_mavg)
    )
  )

rel_pwr_real = pres_long %>%
  dplyr::filter(trial_stage == "NF", block != "1") %>%
  group_by(task, block, trial) %>%
  summarize(
    log_ERS_ipsi_real = mean(log_ERS_ipsi)
    , log_ERS_contra_real = mean(log_ERS_contra)
    )

rel_pwr_compare = left_join(rel_pwr_cal, rel_pwr_real)  

rel_pwr_compare %>%
  ggplot(aes(x = log_ERS_contra_cal, y = log_ERS_contra_real))+
    geom_point()+
    geom_smooth()+
    facet_grid(.~task)+
    ggtitle("NF: Contra Log ERS")+
    xlab("Calculated Log ERS")+
    ylab("Realtime Log ERS")

rel_pwr_compare %>%
  ggplot(aes(x = log_ERS_ipsi_cal, y = log_ERS_ipsi_real))+
    geom_point()+
    geom_smooth()+
    facet_grid(.~task)+
    ggtitle("NF: Ipsi Log ERS")+
    xlab("Calculated Log ERS")+
    ylab("Realtime Log ERS")

rel_pwr_compare %>%
  group_by(task) %>%
  summarize(
    contra =cor(log_ERS_contra_cal, log_ERS_contra_real, use = "complete")
    , ipsi = cor(log_ERS_ipsi_cal, log_ERS_ipsi_real, use = "complete")
    )
```

##### Compare Average ('inter') Log Relative Power

The average ERS calculations are virtually identical. This is expected as neither the online nor the offline average ERS calculations (for intermittent feedback in the 'pres' case) involved the use of a data buffer.

```{r}
rel_pwr_real2 = pres_long %>%
  dplyr::filter(trial_stage == "inter", block != "1") %>%
  group_by(task, block, trial) %>%
  summarize(
    log_ERS_ipsi_real = mean(log_ERS_ipsi)
    , log_ERS_contra_real = mean(log_ERS_contra)
    )

rel_pwr_compare2 = left_join(rel_pwr_cal, rel_pwr_real2)

rel_pwr_compare2 %>%
  ggplot(aes(x = log_ERS_contra_cal, y = log_ERS_contra_real))+
    geom_point()+
    geom_smooth()+
    facet_grid(.~task)+
    ggtitle("Inter: Contra Log ERS")+
    xlab("Calculated Log ERS")+
    ylab("Realtime Log ERS")

rel_pwr_compare2 %>%
  ggplot(aes(x = log_ERS_ipsi_cal, y = log_ERS_ipsi_real))+
    geom_point()+
    geom_smooth()+
    facet_grid(.~task)+
    ggtitle("Inter: Ipsi Log ERS")+
    xlab("Calculated Log ERS")+
    ylab("Realtime Log ERS")

rel_pwr_compare2 %>%
  group_by(task) %>%
  summarize(
    contra =cor(log_ERS_contra_cal, log_ERS_contra_real, use = "complete")
    , ipsi = cor(log_ERS_ipsi_cal, log_ERS_ipsi_real, use = "complete")
    )
```

### Look at Presentation Data

##### Time Stamps

First, we'll examine the 'timestamps'. All iterations were complete, for each trial, and each block, for each block condition. Currently, we can't really evaluate the same for the 'inter' trial stage.

```{r}
pres_long %>%
  filter(block != "1") %>%
  group_by(task, block, trial, trial_stage) %>%
  summarize(
    stage_time = max(iteration)*(1/60)*3
  ) %>%
  ggplot()+
  geom_line(aes(x = trial, y = stage_time))+
  facet_grid(block ~ task + trial_stage)
```


